Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-03-08 paper/gloaguenAppliedNMF.tex
proof read until protocol

These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions. In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}). Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus is generated based on urban recordings mixing many kind of sound sources.

@@

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files with a cumulative duration of 16min01), quiet street (Q, 33 audio files with a cumulative duration of 77min27), noisy street (N, 24 audio files with a cumulative duration of 56min10) and very noisy street (vN, 9 audio files with a cumulative duration of 21min42). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class. The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

\subsection{Transcription of the recordings}\label{part:simScene}
The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database.
This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}. The control of high level parameters can be handle by the user as the sound class presence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior. It allows too the design of a sound mixture from an annotation text file.
As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.

@@

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes. Each sound class is present in multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Senic, Renault Megane and Renault Clio, Dacia Sandero),  for different speeds and gear ratios. In all, 103 car passages have been recorded. The audio samples of the first two cars (Renault Senic and Renault Megane) are included in the \textit{SimScene}'s sound database (50 audio files totally). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recording can be found in \cite{gloaguen_creation_2017}.

\begin{table}[t]
\centering
\caption{$mTIR$ for each sound environment.}
\begin{tabular}{lc}
 & $mTIR$ (dB)\\ \hline
 park & -9.10 ($\pm$ 7.35) \\
 quiet street & 0.88 ($\pm$ 5.92) \\
 noisy street & 6.96 ($\pm$ 5.16) \\
 very noisy street & 15.75 ($\pm$ 9.78) \\ \hline
\end{tabular}
\label{tab:mTIR}
\end{table}

With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure than the recordings. The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate this adjustment, the mean \textit{Traffic Interfering Ratio} is calculated and summarized on Table \ref{tab:mTIR}. It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}. It quantifies the predominance of the traffic component on the 4 sound environments.

@@

To validate its realism, the corpus of transcribed urban sound scenes is submitted to a perceptual test.

\subsection{Perceptual test}
The perceptual test is conducted with a panel of 50 listeners that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a mix of transcribed and recorded scenes. The total number of sound scenes tested is set at 40. The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes. The second half is composed of the same 30-second transcribed scenes. In order to limit the duration of the test and to preserve the concentration of the listeners, each of them listens a subset of 20 sound scenes, which mix recorded and transcribed samples. Furthermore, all the scenes are normalized to the same sound level, chosen at 65 dB, to prevent the listeners from changing the sound level of their speakers.

@@

The test was administered online on the 8 February 2017 and the number of participant has been reached 12 days later. During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards. The participants could also leave a comment on each audio to explain the rating. Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared having no experience in the listening of urban sound mixtures. Figure \ref{fig:boxPlot_test} summarizes the score distributions of all the listeners for the recorded (rec.) and the transcribed (trans.) scenes.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf}
\caption{Box-and whiskers plot of the rating of realism according to the type (recorded /transcribed).}
\label{fig:boxPlot_test}
\end{figure}

The distributions of the notations according to the type are extremely similar. The mean score for the transcribed scenes is even superior to the recorded ones ($m_{trans.}$ = 5.1 ($\pm 1.6$), $m_{rec}$ = 4.9 ($\pm 1.6$)). These results confirm that all the recorded and the transcribed scenes are perceived in a similar way by the panel. More details on the results can be found in  \cite{gloaguen_creation_2017}.

@@

>>>>

These methods are applied on simulated sound scenes in order to compare the estimated sound levels with a non ambiguous reference. In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}). Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus by considering more realistic urban sound environments with many kind of sound sources.

@@

\ml{Il faut faire une figure et un paragraphe resumant la figure. Voici un essai (je pense d'ailleurs que cela devrait faire partie des claims dans l'intro) : In order to validate the above described methods, the evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level. The former calls for the use recordings of urban sound environments, and the latter imposes the use of controled use of sound sources with known properties: onset and offset time location, and volume. Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible. We thus propose to design such evalution corpus based on two corpora. The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of "score" is extracted manually, \textit{i.e.} which sound source is present from this time to this time, at which sound level. The second, called the elemental corpus, is a set of monophonic recordings of sound sources that are occurs in the scenes of the reference. The recordings of the elemental corpus are not taken from the reference corpus. The evaluation corpus is then built by sequencing recordings of the elemental corpus according to the scores of the reference corpus. (Le reste est a reecrire avec les termes proposés)}

More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files with a cumulative duration of 16min01), quiet street (Q, 33 audio files with a cumulative duration of 77min27), noisy street (N, 24 audio files with a cumulative duration of 56min10) and very noisy street (vN, 9 audio files with a cumulative duration of 21min42). Then, each audio file is annotated, annotating the start, end time \ml{and level ?} of each sound event along with its sound class. \ml{cette phrase ne dis probablement pas ce qsue tu veux dire annotation et Transcription sont des termes equivalents, The aim of the annotation phase is next to transcribe the recordings}, in order to obtain simulated sound scenes with the same \ml{ca peut preter a confusion en implicitant une distribution statistique : distribution} of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

\subsection{Generation of the evaluation corpus}\label{part:simScene}
The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database.
This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}. \ml{Pas clair The control of high level parameters can be handled by the user as the sound class occurence, the time between each sample of one sound class, the ratio between the sound level of an event class with the background (i.e the \textit{event background ratio} shorten \textit{ebr})\dots Each parameter is completed with a standard deviation to bring random behavior. It allows too the design of a sound mixture from an annotation text file. A remplacer par ceci: The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user. The latter type is considered in this study.}
As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.

@@

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others\textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots sound classes. Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios. Overall, 103 car passages have been recorded. \ml{ajouter une description sur les problématiques d'overfitting et de generalisation. Cela doit etre un rappel d'un texte ecrit dans la partie nmf. A ce titre, cela pourrait etre une extension de l'etude, on en reparle} The audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recordings can be found in \cite{gloaguen_creation_2017}.

\begin{table}[t]
\centering
\caption{$mTIR$ for each sound environment.\ml{les captions doivent etre self-contained, pas d'acronymes non evident}}
\begin{tabular}{lc}
 & $mTIR$ (dB)\\ \hline
 park & -9.10 ($\pm$ 7.35) \\
 quiet street & 0.88 ($\pm$ 5.92) \\
 noisy street & 6.96 ($\pm$ 5.16) \\
 very noisy street & 15.75 ($\pm$ 9.78) \\ \hline
\end{tabular}
\label{tab:mTIR}
\end{table}

With this built-up database, the \textit{SimScene} software and the annotations of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings. The \textit{ebr} parameter is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate this adjustment, the mean \textit{Traffic Interfering Ratio} is calculated and summarized on Table \ref{tab:mTIR}. It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}. It quantifies the predominance of the traffic component for the 4 types of sound environments.

@@

\subsection{Perceptual test}

To evaluate the level of realism of the corpus of transcribed urban sound scenes, a perceptual test is considered.

The perceptual test is conducted with a panel of 50 subjects that are asked to assess the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of a set of transcribed and recorded scenes. The total number of sound scenes tested is set at 40. The first half includes 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{park}, 6 from \textit{quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the recorded scenes. The second half is composed of the transcribed scenes coresponding to the 20 reference scenes. In order to limit the duration of the test to preserve the concentration of the subjects, each subject listens to only a subset of 20 sound scenes. All the scenes are normalized to the same sound level, chosen at 65 dB, and the subjects are not allowed to change the output sound level once set at the beginning of the experiment.

@@

The test was available online on the 8 February 2017 and the needed number of participant has been reached 12 days later. During the test, the participants had the possibility to listen to each scene as many times as wanted before assessing, without being able to change their judgment afterwards. The participants could also leave a comment on each audio to explain the rating \footnote{The interface of the experiment is available \url{http://soundthings.org/research/xpRealism}}. Based on the information provided, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared having no experience in the listening of urban sound mixtures. Figure \ref{fig:boxPlot_test} summarizes the score distributions of all the listeners for the recorded (rec.) and the transcribed (trans.) scenes.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/testPerceptif_boxplotType_EN.pdf}
\caption{Box and whiskers plot of the rating of realism according to the type (recorded /transcribed). \ml{Il faut tout expliciter, as tu vraiment besoin de toutes ces informations ?}}
\label{fig:boxPlot_test}
\end{figure}

The distributions of the notations according to the type are extremely similar. The mean score for the transcribed scenes is even superior to the recorded ones ($m_{trans.}$ = 5.1 ($\pm 1.6$), $m_{rec}$ = 4.9 ($\pm 1.6$)). These results confirm that the recorded and the transcribed scenes are perceived in a similar way by the panel. More details on the results can be found in  \cite{gloaguen_creation_2017}.

\ml{ce serait quand meme bien d'avoir un test stat, au moins un ttest}

@@

UPTO
